\name{}: Supplement data

### 2021-04-27 paper/main.tex
.

While in this initial formulation of a GCNConv the node-wise update
step is defined by the sum over all neighboring node representations,
we can alter this formulation to other message passing schemes.  We
can rearrange the order of activation function $\sigma$, aggregation
$\mathrm{AGG}$, and linear neural layer $\mathrm{MLP}$ with this
formulation as proposed by \citet{GENConv2020}:
\begin{equation}
	\mathbf{x}_i^{\prime} = \mathrm{MLP} \left( \mathbf{x}_i +
	\mathrm{AGG} \left( \left\{
	\mathrm{\sigma} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
	: j \in \mathcal{N}(i) \right\} \right)
	\right)
\end{equation}
where we only consider
$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote this
generalized layer type as \textsc{GENConv} following the notation of
PyTorch Geometric \cite{PytorchGeometric}.  While the reordering is
mainly important for numerical stability, this alteration also addresses
the vanishing gradient problem for deeper convolutional networks
\cite{GENConv2020}. Additionally, we can also generalize the
aggregation function to allow different weighting functions such as
learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming
signals for each vertex, substituting the averaging step in
\textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
vanishing gradients and signal fading for large scale and highly
connected graphs, each propagation step in \textsc{GENConv} emphasizes
signals with values close to $0$ and $1$. The same convolutional
filter and weight matrix are applied to and learned for all nodes
simultaneously. % , and the resulting information\todo{Which information?
  % Specify} hold no information on their own connectivity.
We further employ another mechanism to avoid redundancy and fading
signals in stacked graph convolutional networks, using residual
connections and a normalization scheme \cite{DeepGCN2019,
  DeeperGCN2020} as shown in Supplemtary Figure 3.  The residual
blocks are reusable and can be stacked multiple times.
% The structure of the GNN architecture we use is shown
% in Figure~\ref{fig:ResGraphConvBlocks}.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. As this is quite common for DTI datasets, we do not
omit but impute the $MicroAUC$ interpolating linearly for those
entities using the accuracy for this subset:
\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Note that drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

Further, we choose the AUROC as primary metric to compare different
methods over other measures such as the area under precision recall
curve (AUPRC); AUPRC is sensitive to data imbalances \cite{Jeni2013}
and therefore more challenging to apply to comparing different DTI
prediction methods.

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|p{1.1cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|}
    \hline
    &\multicolumn{4}{c|}{(a) STITCH results}& \multicolumn{4}{c|}{(b) Yamanishi results}\\
    \hline
    \name{} results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
    &\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
    &Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
    \hline
    MolPred&$0.69$&$0.65$&$0.69$&$0.67$&$0.66$&$0.67$&$0.66$&$0.64$\\
    \hline
    OntoPred&$0.88$&$0.87$&$0.92$&$0.93$&$0.80$&$0.79$&$0.83$&$0.82$\\
    \hline
    \name{} & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
    \hline
  \end{tabular}
  \caption{\label{tab:Results}Results for \name{} on the STITCH
    and Yamanishi datasets evaluated with 5-fold cross-validation. We 
    call the model using only molecular features
    \textit{MolPred} and the model using only ontology-based
    features \textit{OntoPred}. \name{} combines both types of
    features. For a comparison of MicroAUC$_p$ and MicroAUC$'_p$ (See Section \ref{sec:evaluation} for more information) see Supplementary Table 2. }
\end{table}

@@

We find that the model using ontology-based features
(\textit{OntoPred}) is showing better performance on STITCH compared
to using only molecular features. We also observe that only the model
using ontology-based features results in increased performance when
incorporating the PPI graph. This increase can be observed with
different graph neural network architectures and configurations
(Supplementary Table 1). While the {\em
  GCNConv} and {\em GENConv} architecture already shows some minor
improvement, the use of \textit{ResGraphConv} results in larger
performance improvements.  {\em ResGraphConv} blocks add a large
amount of additional learnable parameters to the network, leading to
more expressive power. To test whether the observed improvement is due
to the number of learnable parameters added or the result of better
exploiting the information about PPIs, we experiment with a graph
model in which all graph convolutional neural layers in the residual
blocks are removed, resulting in a model with similar parameters but
without the ability to use graph-based information. This pruned
network, with no information on the protein--protein interactions,
reached very similar results to the original \textit{OntoPred} model
and showed no improvement (Supplementary Table 1).

>>>>

While in this initial formulation of a GCNConv the node-wise update
step is defined by the sum over all neighboring node representations,
we can alter this formulation to other message passing schemes.  We
can rearrange the order of activation function $\sigma$, aggregation
$\mathrm{AGG}$, and linear neural layer $\mathrm{MLP}$ with this
formulation as proposed by \citet{GENConv2020}:
\begin{equation}
	\mathbf{x}_i^{\prime} = \mathrm{MLP} \left( \mathbf{x}_i +
	\mathrm{AGG} \left( \left\{
	\mathrm{\sigma} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
	: j \in \mathcal{N}(i) \right\} \right)
	\right)
\end{equation}
where we only consider
$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote this
generalized layer type as \textsc{GENConv} following the notation of
PyTorch Geometric \cite{PytorchGeometric}.  While the reordering is
mainly important for numerical stability, this alteration also addresses
the vanishing gradient problem for deeper convolutional networks
\cite{GENConv2020}. Additionally, we can also generalize the
aggregation function to allow different weighting functions such as
learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming
signals for each vertex, substituting the averaging step in
\textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
vanishing gradients and signal fading for large scale and highly
connected graphs, each propagation step in \textsc{GENConv} emphasizes
signals with values close to $0$ and $1$. The same convolutional
filter and weight matrix are applied to and learned for all nodes
simultaneously. % , and the resulting information\todo{Which information?
  % Specify} hold no information on their own connectivity.
We further employ another mechanism to avoid redundancy and fading
signals in stacked graph convolutional networks, using residual
connections and a normalization scheme \cite{DeepGCN2019,
  DeeperGCN2020} as shown in Supplementary Figure 3.  The residual
blocks are reusable and can be stacked multiple times.
% The structure of the GNN architecture we use is shown
% in Figure~\ref{fig:ResGraphConvBlocks}.

@@

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. As this is quite common for DTI datasets, we do not
omit but impute the $MicroAUC$ interpolating linearly for those
entities using the accuracy for this subset:
\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).  We further compare
MicroAUC$_p$ with imputation and MicroAUC$'_p$ without imputing
undefined $MicroAUC$ values (but omitting them) in Supplementary Table
1.

Further, we choose the AUROC over other measures such as the area
under precision recall curve (AUPRC) as primary metric to compare
different methods; AUPRC is sensitive to data imbalances
\cite{Jeni2013} and therefore more challenging to apply to comparing
different DTI prediction methods.

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{|p{1.1cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|}
    \hline
    &\multicolumn{4}{c|}{(a) STITCH results}& \multicolumn{4}{c|}{(b) Yamanishi results}\\
    \hline
    \name{} results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
    &\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
    &Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
    \hline
    MolPred&$0.69$&$0.65$&$0.69$&$0.67$&$0.66$&$0.67$&$0.66$&$0.64$\\
    \hline
    OntoPred&$0.88$&$0.87$&$0.92$&$0.93$&$0.80$&$0.79$&$0.83$&$0.82$\\
    \hline
    \name{} & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
    \hline
  \end{tabular}
  \caption{\label{tab:Results}Results for \name{} on the STITCH
    and Yamanishi datasets evaluated with 5-fold cross-validation. We 
    call the model using only molecular features
    \textit{MolPred} and the model using only ontology-based
    features \textit{OntoPred}. \name{} combines both types of
    features.}
\end{table}

@@

We find that the model using ontology-based features
(\textit{OntoPred}) is showing better performance on STITCH compared
to using only molecular features. We also observe that only the model
using ontology-based features results in increased performance when
incorporating the PPI graph. This increase can be observed with
different graph neural network architectures and configurations
(Supplementary Table 2). While the {\em
  GCNConv} and {\em GENConv} architecture already shows some minor
improvement, the use of \textit{ResGraphConv} results in larger
performance improvements.  {\em ResGraphConv} blocks add a large
amount of additional learnable parameters to the network, leading to
more expressive power. To test whether the observed improvement is due
to the number of learnable parameters added or the result of better
exploiting the information about PPIs, we experiment with a graph
model in which all graph convolutional neural layers in the residual
blocks are removed, resulting in a model with similar parameters but
without the ability to use graph-based information. This pruned
network, with no information on the protein--protein interactions,
reached very similar results to the original \textit{OntoPred} model
and showed no improvement (Supplementary Table 2).

### 2021-04-27 paper/supplement/supplement.tex
.

@@

\begin{figure}[ht]%figure1
 	\centerline{\includegraphics[width=0.3\textwidth]{../figures/ResGraphConvBlocks.png}}
 	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
 	\label{fig:ResGraphConvBlocks}
\end{figure}

\begin{figure}[ht]%figure1
	\centerline{\includegraphics[width=1\columnwidth]{../figures/full_model_all_layers.png}}
	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
	\label{fig:FullModelAllLayers}
\end{figure}

\begin{table}[ht]
	\centering

	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{1.2cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
		\hline
		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
		\hline
		MolPred & $0.69$ & $0.67$& $0.67$& $0.68$& $0.65$& $0.69$\\
		\hline
		OntoPred &$0.88$ &$0.87$ &$0.88$&$0.90$&$0.86$&$0.92$\\
		\hline
		\name& $0.89$&$0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
		\hline
	\end{tabular}
	\caption{\label{tab:Results}Results for various neural graph convolutional methods enhancing MolPred, OntoPred and \name{} over the STITCH dataset. We evaluated all methods over different stacking heights without and with residual graph convolution blocks (RGC, ResGraphConv) as depicted in Supplementary figure \ref{fig:ResGraphConvBlocks}. ``Empty RGC`` denotes a RGC block with a plain neural layer substituting the GraphConv layer.}
\end{table}

@@

>>>>

\clearpage

@@

\clearpage
\begin{figure}[ht]%figure1
 	\centerline{\includegraphics[width=0.3\textwidth]{../figures/ResGraphConvBlocks.png}}
 	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
 	\label{fig:ResGraphConvBlocks}
\end{figure}

\clearpage
\begin{figure}[ht]%figure1
	\centerline{\includegraphics[width=1\columnwidth]{../figures/full_model_all_layers.png}}
	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
	\label{fig:FullModelAllLayers}
\end{figure}

\clearpage

@@

\clearpage
\begin{table}[ht]
	\centering

	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{1.2cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
		\hline
		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
		\hline
		MolPred & $0.69$ & $0.67$& $0.67$& $0.68$& $0.65$& $0.69$\\
		\hline
		OntoPred &$0.88$ &$0.87$ &$0.88$&$0.90$&$0.86$&$0.92$\\
		\hline
		\name& $0.89$&$0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
		\hline
	\end{tabular}
	\caption{\label{tab:Results}Results for various neural graph
          convolutional methods enhancing MolPred, OntoPred and
          \name{} over the STITCH dataset. We evaluated all methods
          over different stacking heights without and with residual
          graph convolution blocks (RGC, ResGraphConv, see
          Supplementary figure \ref{fig:ResGraphConvBlocks}). ``Empty
          RGC'' denotes a RGC block with a plain neural layer
          substituting the GraphConv layer.}
\end{table}
